---
title: "Assignment 5"
author: "Amelia Farrell"
date: "October 11th 2021"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir = 'C:/Users/Amelia/Documents/Bellevue/dsc520/data')
survey <- read.csv("student-survey.csv")
library(ggplot2)
library(GGally)
```
## Note - Remove example code and comments before submitting assignment.  Producing a professional R Markdown document is the goal. 

#i. Covariance of Survey variables
## Correlation Coefficients can tell us how two variables are related. This does not tell us that one is influencing another, but rather the state of their relationship (between -1 and 1).
```{r, echo=FALSE}
cor(survey)
```

## The above correlation tells us that there is a strong negative correlation (or relationship) (-0.88) between time spent reading and time spent watching TV. Meaning that as one variable increases the other decreases. There also appears to be a moderate positive correlation between and time spent watching TV and reported happiness (0.63). As well as a low negative correlation between and time spent reading and reported happiness (-0.43). When only looking at these correlations, one may think that watching TV increases happiness. However, we must be aware that anything under 0.5/-0.5 is a low correlation and Correlation Coefficients do not indicate causation. 
## We can also visulize this using ggpairs
```{r, echo=FALSE}
ggpairs(survey)
```



#ii. Examining the Survey data variables
```{r, echo=FALSE}
summary(survey)
```

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## References

