---
title: "ASSIGNMENT 7"
author: "Amelia Farrell"
date: "October 18th 2021"
output:
  word_document: default
  pdf_document: default
---

## Multiple Linear Regression
Making predictions in earning potential

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir = "C:/Users/Amelia/Documents/Bellevue/dsc520/completed/assignment07")
heights_df <- read.csv("heights.csv")
```


# Fit a linear model
```{r, echo=TRUE}
earn_lm <-  lm(earn ~ age + sex + height + ed + race, data = heights_df)
```

Viewing the summary of your model
```{r, echo=FALSE}
summary(earn_lm)

predicted_df <- data.frame(
  earn = predict(earn_lm, heights_df),
  ed=heights_df$ed, race=heights_df$race, height=heights_df$height,
  age=heights_df$age, sex=heights_df$sex
  )
```

Compute deviation (i.e. residuals)
```{r, echo=TRUE}
mean_earn <- mean(heights_df$earn)
```

Corrected Sum of Squares Total
```{r, echo=FALSE}
sst <- sum((mean_earn - heights_df$earn)^2)
```

Corrected Sum of Squares for Model
```{r, echo=TRUE}
ssm <- sum((mean_earn - predicted_df$earn)^2)
```


Residuals
```{r, echo=TRUE}
residuals <- heights_df$earn - predicted_df$earn
```

Sum of Squares for Error
```{r, echo=TRUE}
sse <- sum(residuals^2)
```

R Squared R^2 = SSM/SST
```{r, echo=TRUE}
r_squared <- ssm/sst
r_squared
```

Number of observations
```{r, echo=TRUE}
n <- nrow(heights_df)
n
```

Number of regression parameters
```{r, echo=TRUE}
p <- 2
p
```

Corrected Degrees of Freedom for Model (p-1)
```{r, echo=TRUE}
dfm <- p-1
dfm
```

Degrees of Freedom for Error (n-p)
```{r, echo=TRUE}
dfe <- n-p
dfe
```

Corrected Degrees of Freedom Total:   DFT = n - 1
```{r, echo=TRUE}
dft <- n-1
dft
```

Mean of Squares for Model:   MSM = SSM / DFM
```{r, echo=TRUE}
msm <- ssm/dfm
msm
```

Mean of Squares for Error:   MSE = SSE / DFE
```{r, echo=TRUE}
mse <- sse/dfe
mse
```

Mean of Squares Total:   MST = SST / DFT
```{r, echo=TRUE}
mst <- sst/dft
mst
```

F Statistic F = MSM/MSE
```{r, echo=TRUE}
f_score <- msm/mse
f_score
```

Adjusted R Squared R2 = 1 - (1 - R2)(n - 1) / (n - p)
```{r, echo=TRUE}
adjusted_r_squared <- 1 - (1 - r_squared)*(n - 1) / (n - p)
adjusted_r_squared
```





## References

Field, A., J. Miles, and Z. Field. 2012. Discovering Statistics Using R. SAGE Publications. https://books.google.com/books?id=wd2K2zC3swIC.

Lander, J. P. 2014. R for Everyone: Advanced Analytics and Graphics. Addison-Wesley Data and Analytics Series. Addison-Wesley. https://books.google.com/books?id=3eBVAgAAQBAJ.
