---
title: "11.2 Exercise"
author: "Amelia Farrell"
date: "November 8st 2021"
output:
  word_document: default
  pdf_document: default
---

# Introduction to Machine Learning

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir = "C:/Users/Amelia/Documents/Bellevue/dsc520/completed/Final-Project")
binaryclassifier <- read.csv("binary-classifier-data.csv")
trinaryclassifier <- read.csv("trinary-classifier-data.csv")
library(ggplot2)
library(purrr)
library(dplyr)
```

## K nearest neighbors (Binary and Trinary classifier data)

i. In this section we will be fitting a k nearest neighbors algirithm to two diferent data sets.

Before build our modelds, lets plot the data from each dataset using a scatter plot.
```{r, echo=FALSE}
names(Food_Loss)
```











## References

Field, A., J. Miles, and Z. Field. 2012. Discovering Statistics Using R. SAGE Publications. https://books.google.com/books?id=wd2K2zC3swIC.

Lander, J. P. 2014. R for Everyone: Advanced Analytics and Graphics. Addison-Wesley Data and Analytics Series. Addison-Wesley. https://books.google.com/books?id=3eBVAgAAQBAJ.




